### 2012-09-18 doc/prelim/document/chapters/linear_problem/linear_problem.tex
ufinished parallel matvec section

\begin{figure}[htpb!]
  \begin{center}
    \scalebox{1.75}{
      \input{chapters/linear_problem/orthogonal_residual.pdftex_t} }
  \end{center}
  \caption{Orthogonality constraint of the new residual with respect
    to $\mathcal{L}$.}
  \label{fig:linear_projection_constraint}
\end{figure}

@@

\subsubsection{Parallel Matrix-Vector Multiplications}
\label{subsubsec:parallel_mat_vec_mutliply}
We finally consider parallel matrix-vector multiplication operations
using sparse matrices in a compressed storage format. For these
operations, more complex communication patterns will be required given
that the entire global vector is required in order to compute a single
element of the local product vector. Fortunately, the vast majority of
the global vector components will be multiplied by zero due to the
sparse matrix and therefore much of the vector can be
neglected. Instead we only require data from a handful of other
processes.

\begin{figure}[htpb!]
  \begin{center}
    \scalebox{1.5}{
      \input{chapters/linear_problem/partitioned_matrix.pdftex_t} }
  \end{center}
  \caption{Partitioned linear problem.}
  \label{fig:partitioned_linear_problem}
\end{figure}

\subsubsection{Parallel Perfomance Implications for Projection Methods}
\label{subsubsec:projection_method_performance}
For very large distributed machines, global reduction operations
required at several levels of Krylov algorithms stand to reduce
scalablitity and perfomance. Furthermore, communication between
adjacent domains may also cause a bottleneck as the number of domains
used in simulation grows. The end result is that global data must be
collected and communicated. For scaling improvement, we seek a
reduction in these types of operations.

>>>>

\begin{figure}[htpb!]
  \begin{center}
    \scalebox{1.75}{
      \input{chapters/linear_problem/orthogonal_residual.pdftex_t} }
  \end{center}
  \caption{\textbf{Orthogonality constraint of the new residual with
      respect to $\mathcal{L}$.} \textit{By projecting $\ve{r}_0$ onto
      the constraint subspace, we minimize the new residual by
      removing those components.}}
  \label{fig:linear_projection_constraint}
\end{figure}

@@

\subsubsection{Parallel Matrix-Vector Multiplications}
\label{subsubsec:parallel_mat_vec_mutliply}
We finally consider parallel matrix-vector multiplication operations
using sparse matrices in a compressed storage format. For these
operations, more complex communication patterns will be required given
that the entire global vector is required in order to compute a single
element of the local product vector. Fortunately, the vast majority of
the global vector components will be multiplied by zero due to the
sparsity of the matrix and therefore much of the vector can be
neglected. Instead we only require data from a handful of other
processes. Consider the sparse matrix-vector multiply in
Figure~\ref{fig:partitioned_matvec_multiply} that is partitioned on 3
processors.
\begin{figure}[htpb!]
  \begin{center}
    \scalebox{1.5}{
      \input{chapters/linear_problem/partitioned_matrix.pdftex_t} }
  \end{center}
  \caption{\textbf{Sparse matrix-vector multiply operation partitioned
      on 3 processors.}\textit{Each process owns a set of equations
      that correlates to its physical domain.}}
  \label{fig:partitioned_matvec_multiply}
\end{figure}
Each process owns a set of equations that correlate to the physical
domain of which it has ownership. We can break down the equations
owned by each process in order to devise an efficient scheme for the
multiplication. Consider the portion of the matrix-vector multiply
problem owned by process 1 in
Figure~\ref{fig:partitioned_matvec_multiply}. As shown in
Figure~\ref{fig:matvec_proc_1}, the components of the matrix will be
multiplied by pieces of the vector that are owned by all
processors. 
\begin{figure}[htpb!]
  \begin{center}
    \scalebox{1.5}{
      \input{chapters/linear_problem/matvec_proc_1.pdftex_t} }
  \end{center}
  \caption{\textbf{Components of sparse matrix-vector multiply
      operation owned by process 1.} \textit{The numbers above the
      matrix columns indicate the process that owns the piece of the
      global vector they are acting on. In order to compute its local
      components of the matrix-vector product, process 1 needs its
      matrix elements along with all elements of the global vector
      owned by processes 2 and 3. The piece of the matrix shown is
      $\ve{A}_1$ and it is acting locally on $\ve{x}_1$ to compute the
      local piece of the product, $\ve{y}_1$.}}
  \label{fig:matvec_proc_1}
\end{figure}
For those pieces of the matrix that are owned by process 1 that act on
the vector owned by process 1, we do these multiplications first as no
communication is required. Next, process one gathers the components of
the global vector that it requires to complete its part of the vector
product. For this example, the components of matrix that will operate
on the global vector are zero, and therefore no vector elements are
required to be scattered from process 3 to process 1. Those matrix
elements that will act on the piece of the vector owned by process 2
are not all non-zero, and therefore we must gather the entire process
2 vector components onto process 1 to complete the multiplication.
Conversely, processes 2 and 3 must scatter its vector components that
are required by other processes (such as process 1) in order to
complete their pieces of the product. This then implies that these
domain connections for proper gather and scatter combinations must be
constructed a priori. These data structures are typically generated by
a data partitioning library. Mathematically, if we are performing a
global matrix-vector multiply of the form $\ve{A}\ve{x} = \ve{y}$,
then for this example on process 1 we have a sequence of local
matrix-vector multiplications: $\ve{A}_1\ve{x}_1 + \ve{A}_1\ve{x}_2 =
\ve{y}_1$. Here, some of the data is intrinisically local, and some
must be gathered from other processes using the partitioning data
structures.

\subsubsection{Parallel Perfomance Implications for Krylov Methods}
\label{subsubsec:projection_method_performance}
Knowing the parallel characteristics of the key operations we must
perform in order to implement Krylov methods, we can make a few
statements about parallel perfomance and implications for operation on
machines of increasing size.  For very large distributed machines,
global reduction operations required at several levels of Krylov
algorithms stand to reduce scalablitity and perfomance. Furthermore,
communication between adjacent domains may also cause a bottleneck as
the number of domains used in simulation grows. The end result is that
global data must be collected and communicated. For scaling
improvement, we seek a reduction in these types of operations.
